Gap Years in Australia: Effects on Multidimensional Life Satisfaction
========================================================
<a name="top"></a>

Contents:

1. [Data Munging](#SQL)
2. [Cohort Sequence Design](#cohort)
3. [Growth Curve: General Life Satisfaction](#lifeSat)
4. [Growth Curve: Satisfaction with Career Prospects](#careerPros)
5. [Growth Curve: Satisfaction with Future Prospects](#futPros)

<a name="SQL"></a>
SQL Data Munging
--------------
Initial data-munging was done using a series of SQL queries. For direct further education entrants we took those who indicated that they had finished year 12 the year before and indicated that they were currently undertaking some form of teriary study. Gap year students were derived by taking those who had completed year 12 the year before and indicated that they had defered further education and/or confirmed the following year that they had defered further education in the year previous.

Further education drop-outs were taking by exploring if in any of the years under investigation they indicated they had withdrawn from their teriary education course. Gap-year returners composed those who had, at any time during the period of investigation indicated that they currently undertaking teriary education. SQL queries underlying theis subsetting are in .Rmd file associated with this file.
```{r munging1, echo=FALSE}
library(RSQLite)
setwd("/Users/phparker/Dropbox/Databases/LSAY_DATSETS/")
LSAY <- dbConnect(SQLite(), dbname="LSAY.sqlite")
dbListTables(LSAY)
###################
#File Descriptions#
###################
meta <- dbGetQuery(LSAY, "SELECT * FROM META2003")
#####################
#University Entrants#
#####################
#Looking for only those who commence university directly after leaving year12 schooling
########################################################################################
#Need temp files to switch datasets in next section
UniversityEntrantTemp<- dbGetQuery(LSAY, "SELECT * FROM Student2003 where
                               (xcsl2003=1 AND xbac2004=1) OR
                               (xcsl2004=1 AND xbac2005=1) OR
                               (xcsl2005=1 AND xbac2006=1) 
                               ")
######################
#University Deferment#
######################
#Looking for those who graduated year 12 and then confirmed the following year that they had defered study
##########################################################################################################
DeferTemp<- dbGetQuery(LSAY, "SELECT * FROM Student2003 where 
                   (xcsl2003=1 AND (lbca16=1 OR lbca20=1)) OR
                   (xcsl2004=1 AND (lcca22=1 OR lcca26=1)) OR
                   (xcsl2005=1 AND (ldca21a=1 OR ldca25a=1)) OR
                   (xcsl2006=1 AND (leca021a=1 OR leca025a=1))
                   ")
DeferTemp$lbca05temp <-rep(0, nrow(DeferTemp))
############################################################################################
#DEFER: Drop those in University entrants as miscoded
Defer<- DeferTemp[which(DeferTemp$id %in% UniversityEntrantTemp$id ==FALSE) ,]
#UNIVERSITY: Drop those in Defer as miscoded
UniversityEntrant <- UniversityEntrantTemp[which(UniversityEntrantTemp$id %in% DeferTemp$id ==FALSE) ,]
 rm(DeferTemp)
 rm(UniversityEntrantTemp)
#Subgroups
#University drop outs
UniversityEntrant$dropOut <- ifelse(UniversityEntrant$xbac2005==4|
            UniversityEntrant$xbac2006==4|
            UniversityEntrant$xbac2007==4|
            UniversityEntrant$xbac2008==4|
            UniversityEntrant$xbac2009==4|
            UniversityEntrant$xbac2010==4,
            1,0)
#Gap Year Returners
Defer$returners <- ifelse(Defer$xbac2005<4|
            Defer$xbac2006<4|
            Defer$xbac2007<4|
            Defer$xbac2008<4|
            Defer$xbac2009<4|
            Defer$xbac2010<4,
            1,0)
```

The number of adolescence who directly entered university after high-school was ***`r nrow(UniversityEntrant)`***. The number of adolescents that took a gap-year (offered further education but defered) was ***`r nrow(Defer)`***. Of the direct entrants, those who withdraw from further education in the direct entrant group were ***`r table(UniversityEntrant$dropOut)[2]`***. For ther gap-year group, those who eventually returned to further education were ***`r table(Defer$returners)[2]`***.

[top](#top)

<a name="cohort"></a>
Cohort Sequence Design
-----------------------
The PISA database consistst on individuals who are the same age but may be in a veriety of year in school grades. In the current research we wanted to compare individuals in terms of years since high-school graduation rather than age in years. As such, we needed to rearrange the data.

The Satisfaction Items are (How satisfied are you with):

* WORK The work you do, at school, at home or in a job
* LEISURE What you do in your spare time
* GET.ALONG How you get on with people in general
* WAGE The money you get each week
* SOCIAL Your social life
* INDEPENDENCE Your independence - being able to do what you want
* **CAREER.PROS** Your career prospects
* **FUTURE.PROS** Your future
* HOME.LIFE Your life at home
* LIVE.STAND Your standard of living
* RESIDENCE Where you live
* **LIFE.SAT** Your life as a whole

Items in **bold** are the ones used in the analysis.
```{r munging2, echo=FALSE}
UniversityEntrant$sinceHS <- with(UniversityEntrant, ifelse(xcsl2003==1,0,
                                                            ifelse(xcsl2004==1,1,
                                                              ifelse(xcsl2005==1,2,
                                                                ifelse(xcsl2006==1,3,
                                                                  ifelse(xcsl2007==1,4,
                                                                    ifelse(xcsl2008==1,5,
                                                                      ifelse(xcsl2009==1,6,NA)
                                                                        )
                                                                      )
                                                                    )
                                                                  )
                                                                )
                                                            )
                                  )


Defer$sinceHS <- with(Defer, ifelse(xhsl2004==1,0,
                                                            ifelse(xhsl2005==1,1,
                                                              ifelse(xhsl2006==1,2,
                                                                ifelse(xhsl2007==1,3,
                                                                  ifelse(xhsl2008==1,4,
                                                                    ifelse(xhsl2009==1,5,NA)
                                                                        )
                                                                      )
                                                                    )
                                                                  )
                                                                )
                                  )

#Extracting and relabeling life satisfaction
#University
Unames <- c(grep("bj003|cj002|dj010|ej005|fj003|gj010|hj001", names(UniversityEntrant), value=TRUE), 
                 "sinceHS", "dropOut", "id", "sex", "indig", 
              "xbac2004","xbac2005","xbac2006","xbac2007","xbac2008","xbac2009","xbac2010")
                  #"lbca24","lcca30","ldca30","leca030","lfca020","lgca020","lhca038")
UniversityEntrant<- UniversityEntrant[,Unames]
names(UniversityEntrant)[grep("lcj002k|lcj002l", names(UniversityEntrant))] <-c("lcj002m","lcj002n")
names(UniversityEntrant)[grep("ldj010k|ldj010l", names(UniversityEntrant))] <-c("ldj010m","ldj010n")
#Recode unsure/missing
library(car)
UniversityEntrant[,grep("bj003|cj002|dj010|ej005|fj003|gj010|hj001", 
                        names(UniversityEntrant))] <- 
  apply(UniversityEntrant[,grep("bj003|cj002|dj010|ej005|fj003|gj010|hj001",
                                names(UniversityEntrant))],
                                2, function(x) x<- recode(x, "5=NA;1=4;2=3;3=2;4=1"))
#Organise cohort sequence
UniversityEntrantSplit <- split(UniversityEntrant, UniversityEntrant$sinceHS)
for (i in 0:(length(UniversityEntrantSplit)-1)){
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)a$", 
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "WORK")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)b$", 
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "LEISURE")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)c$",
                  names(UniversityEntrant))] <-paste0("W", 1:7-i, "GET.ALONG")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)d$",
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "WAGE")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)e$", 
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "SOCIAL")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)f$", 
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "INDEPENDENCE")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)g$", 
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "CAREER.PROS")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)h$", 
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "FUTURE.PROS")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)i$", 
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "HOME.LIFE")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)j$", 
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "LIVE.STAND")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)m$", 
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "RESIDENCE")
  names(UniversityEntrantSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)n$",
                  names(UniversityEntrant))] <-paste0("W", (1:7)-i, "LIFE.SAT")
  names(UniversityEntrantSplit[[i+1]])[grep("^xbac20(04|05|06|07|08|09|10)$",
                 names(UniversityEntrant))] <-paste0("W", (1:7)-i, "STATUS")
  }
library(plyr)
UniversityEntrant2 <- rbind.fill(lapply(UniversityEntrantSplit,function(y){as.data.frame(y)}))
###############################
#Defer Group
Dnames <- c(grep("bj003|cj002|dj010|ej005|fj003|gj010|hj001", names(Defer), value=TRUE), 
                 "sinceHS", "returners", "id", "sex", "indig",
                "lbca05temp",  "lcca05", "ldca05", "leca005", "lfca005", "lgca005", "lhca005")
Defer<- Defer[,Dnames]
names(Defer)[grep("lcj002k|lcj002l", names(Defer))] <-c("lcj002m","lcj002n")
names(Defer)[grep("ldj010k|ldj010l", names(Defer))] <-c("ldj010m","ldj010n")
#Recode unsure/missing
Defer[,grep("bj003|cj002|dj010|ej005|fj003|gj010|hj001", 
                        names(Defer))] <- 
  apply(Defer[,grep("bj003|cj002|dj010|ej005|fj003|gj010|hj001", names(Defer))],
                                                            2, function(x) x<- recode(x, 
                                                                                      "5=NA;1=4;2=3;3=2;4=1"))
#Organise cohort sequence
DeferSplit <- split(Defer, Defer$sinceHS)
for (i in 0:(length(DeferSplit)-1)){
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)a$", 
                  names(Defer))] <-paste0("W", (1:7)-i, "WORK")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)b$", 
                  names(Defer))] <-paste0("W", (1:7)-i, "LEISURE")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)c$",
                  names(Defer))] <-paste0("W", 1:7-i, "GET.ALONG")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)d$",
                  names(Defer))] <-paste0("W", (1:7)-i, "WAGE")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)e$", 
                  names(Defer))] <-paste0("W", (1:7)-i, "SOCIAL")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)f$", 
                  names(Defer))] <-paste0("W", (1:7)-i, "INDEPENDENCE")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)g$", 
                  names(Defer))] <-paste0("W", (1:7)-i, "CAREER.PROS")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)h$", 
                  names(Defer))] <-paste0("W", (1:7)-i, "FUTURE.PROS")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)i$", 
                  names(Defer))] <-paste0("W", (1:7)-i, "HOME.LIFE")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)j$", 
                  names(Defer))] <-paste0("W", (1:7)-i, "LIVE.STAND")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)m$", 
                  names(Defer))] <-paste0("W", (1:7)-i, "RESIDENCE")
  names(DeferSplit[[i+1]])[grep("^l(bj003|cj002|dj010|ej005|fj003|gj010|hj001)n$",
                  names(Defer))] <-paste0("W", (1:7)-i, "LIFE.SAT")
  names(DeferSplit[[i+1]])[grep("lbca05temp|lcca05|ldca05|leca005|lfca005|lgca005|lhca005",
                  names(Defer))] <-paste0("W", (1:7)-i, "STATUS")
  }
library(plyr)
Defer2 <- rbind.fill(lapply(DeferSplit,function(y){as.data.frame(y)}))
######
```

```{r helperFun, echo=FALSE}
FunPlot_Multi <- function(model=fitQG, data=lifeSat, main="Life Satisfaction",
                          var="LIFE.SAT", paras){
  #Observed Means
  MGS1 <- apply(data[data$group==-1,1:6], 2, mean, na.rm=TRUE)
  MGS2 <- apply(data[data$group==1,1:6], 2, mean, na.rm=TRUE)
  #Observed Trend
  plot(0:5, MGS1[c(6,1:5)], type="l", lty=2, col="blue",
       ylab="Growth", xlab="Years Since Graduation", main=main,
       ylim=c(min(c(MGS1, MGS2))-.01, max(c(MGS1, MGS2))+.01)
      )
  f1 <- function(x) {(paras[1]+paras[4]*-1) + (paras[2]*x+paras[5]*x*-1) + 
                       (paras[3]*x^2+paras[6]*x^2*-1)}
  curve(f1, 0,5,col="blue", add=TRUE)
  f2 <- function(x) {(paras[1]+paras[4]*1) + (paras[2]*x+paras[5]*x*1) + 
                       (paras[3]*x^2+paras[6]*x^2*1)}
  curve(f2, 0,5,col="red", add=TRUE)
  points(0:5, MGS2[c(6,1:5)], col="red", type="l", lty=2)
}

FunPlot_Multi2 <- function(model=fitQG, data=lifeSat, main="Life Satisfaction",
                          var="LIFE.SAT", paras){
  #Observed Means
  MGS1 <- apply(data[data$group.x==0,grep(paste0("W", "[0-5]",var), names(data))], 2, weighted.mean, weights=weights, na.rm=TRUE)
  MGS2 <- apply(data[data$group.x==1,grep(paste0("W", "[0-5]",var), names(data))], 2, weighted.mean, weights=weights, na.rm=TRUE)  #Observed Trend
  plot(0:5, MGS1[c(6,1:5)], type="l", lty=2, col="blue",
       ylab="Growth", xlab="Years Since Graduation", main=main,
       ylim=c(min(c(MGS1, MGS2))-.01, max(c(MGS1, MGS2))+.01)
      )
  f1 <- function(x) {(paras[1]+paras[4]*0) + (paras[2]*x+paras[5]*x*0) + 
                       (paras[3]*x^2+paras[6]*x^2*0)}
  curve(f1, 0,5,col="blue", add=TRUE)
  f2 <- function(x) {(paras[1]+paras[4]*1) + (paras[2]*x+paras[5]*x*1) + 
                       (paras[3]*x^2+paras[6]*x^2*1)}
  curve(f2, 0,5,col="red", add=TRUE)
  points(0:5, MGS2[c(6,1:5)], col="red", type="l", lty=2)
}

Model_Fun <- function(name="LIFE.SAT", 
                      params = list(i=TRUE,s=TRUE,q=TRUE,ig=TRUE,sg=TRUE,qg=TRUE)){
  i = ' i =~ 1*W0temp+1*W1temp+1*W2temp+1*W3temp+1*W4temp+1*W5temp'
  s = 's =~ 0*W0temp+1*W1temp+2*W2temp+3*W3temp+4*W4temp+5*W5temp
  s ~ b*1'
  q = 'q =~ 0*W0temp+1*W1temp+4*W2temp+9*W3temp+16*W4temp+25*W5temp
  q ~ c*1'
  ig = 'i ~ group'
  sg = 's ~ g1*group'
  qg = 'q ~ g2*group
  tpG1 := -1*(b - g1)/(2 * (c - g2))
  tpG2 := -1*(b + g1)/(2 * (c + g2))'
  lines <- c(i,s,q,ig,sg,qg)
  out <- paste(lines[do.call(c, params)], collapse="\n")
  out <- gsub("temp", name, out)
  out
}
```
[top](#top)

<a name="lifeSat"></a>
CFA Growth - Life Satisfaction
-----------------------

Group Parameters significant for Life Satisfaction so this model is plotted at the end of this section.

```{r lifeSat, echo=FALSE, warning=FALSE, message=FALSE}
library(lavaan)
lifeSat<-rbind(UniversityEntrant2[,grep("W[0-5]LIFE.SAT", names(UniversityEntrant2))],
              Defer2[, grep("W[0-5]LIFE.SAT", names(Defer2))])
lifeSat$group <- c(rep(-1, nrow(UniversityEntrant2)), rep(1, nrow(Defer2)))
#Linear
m_linear <- Model_Fun(name="LIFE.SAT", params = list(i=TRUE,s=TRUE,q=F,ig=F,sg=F,qg=F))
fitL <- growth(m_linear, data=lifeSat, missing="fiml")
#Quadratic
m_quadratic <- Model_Fun(name="LIFE.SAT", params = list(i=T,s=T,q=T,ig=F,sg=F,qg=F))
fitQ <- growth(m_quadratic, data=lifeSat, missing="fiml")
Fit <- rbind(fitMeasures(fitL)[c('chisq','df',"cfi",'tli', 'rmsea')], 
             fitMeasures(fitQ)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
anova(fitL, fitQ)
summary(fitQ, standardized=TRUE)
#By Group
mG_quadratic <- Model_Fun(name="LIFE.SAT", params = list(i=T,s=T,q=T,ig=T,sg=T,qg=T))
fitQG <- growth(mG_quadratic, data=lifeSat, missing="fiml")
Fit <- rbind(Fit,  fitMeasures(fitQG)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
summary(fitQG, standardized=TRUE)
##plot
cat("Percentage Missing:\n")
apply(lifeSat[lifeSat$group==-1,1:6], 2, function(x) sum(is.na(x))/length(x))
apply(lifeSat[lifeSat$group==1,1:6], 2, function(x) sum(is.na(x))/length(x))
paras<- parameterEstimates(fitQG)[c(44,13,20,21:23),"est"]
FunPlot_Multi(model=fitQG, data=lifeSat, main="Life Satisfaction", paras)
```
[top](#top)

<a name="careerPros"></a>
CFA Growth - Career Prospects
-----------------------
Group Parameters not significant for Life Satisfaction so pooled model is plotted at the end of this section.

```{r careerPros, echo=FALSE, warning=FALSE, message=FALSE}
careerPros<-rbind(UniversityEntrant2[,grep("W[0-5]CAREER.PROS", names(UniversityEntrant2))],
              Defer2[, grep("W[0-5]CAREER.PROS", names(Defer2))])
careerPros$group <- c(rep(-1, nrow(UniversityEntrant2)), rep(1, nrow(Defer2)))
#Linear
m_linear <- Model_Fun(name="CAREER.PROS", params = list(i=T,s=T,q=F,ig=F,sg=F,qg=F))
fitL <- growth(m_linear, data=careerPros, missing="fiml")
#Quadratic
m_quadratic <- Model_Fun(name="CAREER.PROS", params = list(i=T,s=T,q=T,ig=F,sg=F,qg=F))
fitQ <- growth(m_quadratic, data=careerPros, missing="fiml")
Fit <- rbind(fitMeasures(fitL)[c('chisq','df',"cfi",'tli', 'rmsea')], 
             fitMeasures(fitQ)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
anova(fitL, fitQ)
summary(fitQ, standardized=TRUE)
#By Group
mG_quadratic <- Model_Fun(name="CAREER.PROS", params = list(i=T,s=T,q=T,ig=T,sg=T,qg=T))
fitQG <- growth(mG_quadratic, data=careerPros, missing="fiml")
Fit <- rbind(Fit,  fitMeasures(fitQG)[c('chisq','df',"cfi",'tli', 'rmsea')])
summary(fitQG, standardized=TRUE)
cat("Percentage Missing:\n")
apply(careerPros[careerPros$group==-1,1:6], 2, function(x) sum(is.na(x))/length(x))
apply(careerPros[careerPros$group==1,1:6], 2, function(x) sum(is.na(x))/length(x))
paras<- parameterEstimates(fitQG)[c(44,13,20,21:23),"est"]
FunPlot_Multi(model=fitQG, data=careerPros, main="Career Prospects", paras)
```
[top](#top)

<a name="futPros"></a>
CFA Growth - Future Prospects
-----------------------
Group Parameters not significant for Life Satisfaction so pooled model is plotted at the end of this section.

```{r futurePros, echo=FALSE, warning=FALSE, message=FALSE}
futPros<-rbind(UniversityEntrant2[,grep("W[0-5]FUTURE.PROS", names(UniversityEntrant2))],
              Defer2[, grep("W[0-5]FUTURE.PROS", names(Defer2))])
futPros$group <- c(rep(-1, nrow(UniversityEntrant2)), rep(1, nrow(Defer2)))
#Linear
m_linear <- Model_Fun(name="FUTURE.PROS", params = list(i=T,s=T,q=F,ig=F,sg=F,qg=F))
fitL <- growth(m_linear, data=futPros, missing="fiml")
#Quadratic
m_quadratic <- Model_Fun(name="FUTURE.PROS", params = list(i=T,s=T,q=T,ig=F,sg=F,qg=F))
fitQ <- growth(m_quadratic, data=futPros, missing="fiml")
Fit <- rbind(fitMeasures(fitL)[c('chisq','df',"cfi",'tli', 'rmsea')], 
             fitMeasures(fitQ)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
anova(fitL, fitQ)
summary(fitQ)
#By Group
mG_quadratic <- Model_Fun(name="FUTURE.PROS", params = list(i=T,s=T,q=T,ig=T,sg=T,qg=T))
fitQG <- growth(mG_quadratic, data=futPros, missing="fiml")
Fit <- rbind(Fit,  fitMeasures(fitQG)[c('chisq','df',"cfi",'tli', 'rmsea')])
summary(fitQ, standardized=TRUE)
cat("Percentage Missing:\n")
apply(futPros[futPros$group==-1,1:5], 2, function(x) sum(is.na(x))/length(x))
apply(futPros[futPros$group==1,1:5], 2, function(x) sum(is.na(x))/length(x))
paras<- parameterEstimates(fitQG)[c(44,13,20,21:23),"est"]
FunPlot_Multi(model=fitQG, data=futPros, main="Future Prospects", paras)
```
[top](#top)

Extract Matching Variables
----------------------------
```{r matchVar, echo=TRUE}
#Single item key|value pairs
baseData <- dbGetQuery(LSAY, "SELECT * FROM Student2003")
WaveAdict_singleitem <- list(id = 'id', stateid = 'State territory', schoolid = 'School ID',
                             loc = 'location class', sex = 'gender',
                             sisced = 'education expectations', indig = 'indigenous status', 
                             laa005 = 'ed plans',
                             laa006 = 'parent ed plaan', escs = "ses index from pisa results",
                             laa007 = 'friend ed plan',
                             laa025 = 'social comparison english',
                             laa026 = 'social comparison english', 
                             laa027 = 'social comparison total',
                             xcsl2003 = 'grade', lad001 = 'work',
                             xath2003 = 'live parents',
                             belong = 'sense of belonging', intmat = 'interest in math',
                             instmot = 'intrinsic motivation', matheff = 'self-efficacy',
                             anxmat = 'anxiety', scmat = 'self-concept', complrn = 'competative',
                             cooplrn = 'cooperative', teachsup = 'teacher support',
                            disclim = 'discipline climate')

#Multi item scal and index item key|value pairs    
WaveAdict_multiitem<- list(pv1 = "plausible values for achievement - First",
                           HRS = "hours how spent per week - 6 item"
                           #st30q = "Attitudes toward education career - 8 item" 
                           )

##Print Scale descriptions and items flags
print("Single Item scales")
for (i in seq_along(WaveAdict_singleitem)){
  cat(paste(names(WaveAdict_singleitem)[i], WaveAdict_singleitem[[i]][1], sep=":\t"), sep="\n")
}
print("Multi Item scales")
for (i in seq_along(WaveAdict_multiitem)){
  cat(paste(names(WaveAdict_multiitem)[i], WaveAdict_multiitem[[i]][1], sep=":\t"), sep="\n")
}
#extraction of multi item locations
baseLocations <- paste0(paste0("^(", paste(names(WaveAdict_multiitem),collapse="|"), ")", ".*$"))
FeatureIndex<-grep(baseLocations, names(baseData), value=TRUE)
#Analaysis database is reduction from full lsay data
matchData <- data.frame(baseData[,names(WaveAdict_singleitem)], baseData[, FeatureIndex])
#recode missing
library(car)
matchData[,-c(1)] <- data.frame(apply(matchData[,-c(1)], 2,
                                      function(x) recode(x, "999=NA; 998=NA")))
matchData[,-c(1)] <- data.frame(apply(matchData[,-c(1)], 2,
                              function(x) recode(x, "999=NA; 998=NA")))
#Amount of missing data
apply(matchData,2, function(x) sum(is.na(x))/nrow(matchData))
#na.omit small amount of missing data for matching
data <- rbind(UniversityEntrant2[,names(UniversityEntrant2)%in%names(Defer2)],
              Defer2[,names(Defer2)%in%names(UniversityEntrant2)])
data$group <- c(rep(0, nrow(UniversityEntrant2)), rep(1, nrow(Defer2)))
matchData <- merge(matchData, data[, c("id", "group")], by="id", by.all=FALSE)
matchData <- na.omit(matchData)
```
[top](#top)

Propensity score matching
-----------------------------
```{r PSM, echo=TRUE, cache=TRUE}
require(ggplot2); require(MatchIt)
meandifftable<- function (x){
  post<-data.frame(x$sum.matched[4])
  matchID <- as.vector (row.names (post) )
  names(post)[1]<-c("m_mean_diff")
  post$absolute<- abs(post[1])
  total2<-post[order (-post$absolute, na.last=NA) ,]
  meandiffover1<- subset(total2[1], total2[1]> .1 | total2[1]< -.1)
  meandiffover1
}

all_meandiffplot <- function (x) {
  adiff<-data.frame(x$sum.all)
  names(adiff)[4]<-c("all_mean_diff")
  diffplot<-ggplot(adiff, aes(all_mean_diff) )
  diffplot<- diffplot+ geom_histogram (fill="grey")
  diffplot<- diffplot+ geom_density (colour="red")
  diffplot<-diffplot+xlim(-.5, .5)
  diffplot
}

matched_meandiffplot <- function (x) {
  mdiff<-data.frame(x$sum.matched)
  names(mdiff)[4]<-c("matched_mean_diff")
  diffplot<-ggplot(mdiff, aes(matched_mean_diff) )
  diffplot<- diffplot+ geom_histogram (fill="grey")
  diffplot<- diffplot+ geom_density (colour="red")
  diffplot<-diffplot+xlim(-.5, .5)
  diffplot
}

all_meandiffcount<-function (x){
  all<-data.frame(x$sum.all[4])
  all$all_group[all[1] > .25]<- "Large"
  all$all_group[all[1] < -.25] <- "Large"
  all$all_group[all[1] > .20 & all[1] < .25 ] <- "Medium"
  all$all_group[all[1] < -.20 & all[1] > -.25] <- "Medium"
  all$all_group[all[1] < .20 & all[1] > .00]<- "Small"
  all$all_group[all[1] > -.20 & all[1] < .00] <- "Small"
  table(all$all_group)
}

matched_meandiffcount<-function (x){
  matched<-data.frame(x$sum.matched[4])
  matched$matched_group[matched[1] > .25]<- "Large"
  matched$matched_group[matched[1] < -.25] <- "Large"
  matched$matched_group[matched[1] > .20 & matched[1] < .25 ] <- "Medium"
  matched$matched_group[matched[1] < -.20 & matched[1] > -.25] <- "Medium"
  matched$matched_group[matched[1] < .20 & matched[1] >= .00]<- "Small"
  matched$matched_group[matched[1] > -.20 & matched[1] <= .00] <- "Small"
  table(matched$matched_group)
}

matchData[,c("stateid", "schoolid", "loc")] <- apply(matchData[,c("stateid", "schoolid", "loc")],2, factor)
#drop location and school and state for now
set.seed(102)
forM <- paste("group ~ ", paste0(names(matchData)[-c(1:4, 18, 36)],collapse=" + ") )
#forM1 <- paste(forM, " + I(anxmat*disclim)")
M1 <- matchit(as.formula(forM), data=matchData, method= "nearest", discard="both",
              caliper = .20, ratio=3, distance="logit", distance.options=list(maxit=100))
SM1 <- summary(M1,  addlvariables=NULL, interactions=TRUE,standardize=TRUE)
head(meandifftable(SM1),5)
par(mfrow=c(1,2))
all_meandiffplot(SM1)
matched_meandiffplot(SM1)
par(mfrow=c(1,1))
all_meandiffcount(SM1)
matched_meandiffcount(SM1)
#save matched data with weights
dmatch <- match.data(M1)
nrow(dmatch)
with(dmatch, table(group))

dmatch <- merge(data, dmatch, by="id", by.all=FALSE)
```


Matched Life Satisfaction
-------------------------------
Adding Weights

```{r MlifeSat, echo=TRUE, cache=TRUE}
library(lavaan.survey)
#Linear
m_linear <- Model_Fun(name="LIFE.SAT", params = list(i=TRUE,s=TRUE,q=F,ig=F,sg=F,qg=F))
des <- svydesign(ids = ~id, weights = ~weights, data = dmatch)
fitL <- growth(m_linear, data=dmatch, missing="fiml")
fitL <- lavaan.survey(fitL, survey.design = des, estimator="ML")
#Quadratic
m_quadratic <- Model_Fun(name="LIFE.SAT", params = list(i=T,s=T,q=T,ig=F,sg=F,qg=F))
fitQ <- growth(m_quadratic, data=dmatch, missing="fiml")
fitQ <- lavaan.survey(fitQ, survey.design = des, estimator="ML")
fitQ <- growth(m_quadratic, data=dmatch, missing="fiml")
fitQ <- lavaan.survey(fitQ, survey.design = des, estimator="ML")
Fit <- rbind(fitMeasures(fitL)[c('chisq','df',"cfi",'tli', 'rmsea')], 
             fitMeasures(fitQ)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
anova(fitL, fitQ)
summary(fitQ, standardized=TRUE)
#By Group
mG_quadratic <- Model_Fun(name="LIFE.SAT", params = list(i=T,s=T,q=T,ig=T,sg=T,qg=T))
mG_quadratic <- gsub("group", "group.x", mG_quadratic)
fitQG <- growth(mG_quadratic, data=dmatch, missing="fiml")
fitQG <- lavaan.survey(fitQG, survey.design = des, estimator="ML")
Fit <- rbind(Fit,  fitMeasures(fitQG)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
summary(fitQG, standardized=TRUE)
##plot
paras<- parameterEstimates(fitQG)[c(44,13,20,21:23),"est"]
FunPlot_Multi2(model=fitQG, data=dmatch, main="Life Satisfaction",var="LIFE.SAT", paras)
```

Matched Career Prospects
-------------------------------
Adding Weights

```{r McareerPros, echo=TRUE, cache=TRUE}
library(lavaan.survey)
#Linear
m_linear <- Model_Fun(name="CAREER.PROS", params = list(i=TRUE,s=TRUE,q=F,ig=F,sg=F,qg=F))
des <- svydesign(ids = ~id, weights = ~weights, data = dmatch)
fitL <- growth(m_linear, data=dmatch, missing="fiml")
fitL <- lavaan.survey(fitL, survey.design = des, estimator="ML")
#Quadratic
m_quadratic <- Model_Fun(name="CAREER.PROS", params = list(i=T,s=T,q=T,ig=F,sg=F,qg=F))
fitQ <- growth(m_quadratic, data=dmatch, missing="fiml")
fitQ <- lavaan.survey(fitQ, survey.design = des, estimator="ML")
fitQ <- growth(m_quadratic, data=dmatch, missing="fiml")
fitQ <- lavaan.survey(fitQ, survey.design = des, estimator="ML")
Fit <- rbind(fitMeasures(fitL)[c('chisq','df',"cfi",'tli', 'rmsea')], 
             fitMeasures(fitQ)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
anova(fitL, fitQ)
summary(fitQ, standardized=TRUE)
#By Group
mG_quadratic <- Model_Fun(name="CAREER.PROS", params = list(i=T,s=T,q=T,ig=T,sg=T,qg=T))
mG_quadratic <- gsub("group", "group.x", mG_quadratic)
fitQG <- growth(mG_quadratic, data=dmatch, missing="fiml")
fitQG <- lavaan.survey(fitQG, survey.design = des, estimator="ML")
Fit <- rbind(Fit,  fitMeasures(fitQG)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
summary(fitQG, standardized=TRUE)
##plot
paras<- parameterEstimates(fitQG)[c(44,13,20,21:23),"est"]
FunPlot_Multi2(model=fitQG, data=dmatch, main="Career Prospects",var="CAREER.PROS", paras)
```

Matched Future Prospects
-------------------------------
Adding Weights

```{r MfuturePros, echo=TRUE, cache=TRUE}
library(lavaan.survey)
#Linear
m_linear <- Model_Fun(name="FUTURE.PROSS", params = list(i=TRUE,s=TRUE,q=F,ig=F,sg=F,qg=F))
des <- svydesign(ids = ~id, weights = ~weights, data = dmatch)
fitL <- growth(m_linear, data=dmatch, missing="fiml")
fitL <- lavaan.survey(fitL, survey.design = des, estimator="ML")
#Quadratic
m_quadratic <- Model_Fun(name="FUTURE.PROS", params = list(i=T,s=T,q=T,ig=F,sg=F,qg=F))
fitQ <- growth(m_quadratic, data=dmatch, missing="fiml")
fitQ <- lavaan.survey(fitQ, survey.design = des, estimator="ML")
fitQ <- growth(m_quadratic, data=dmatch, missing="fiml")
fitQ <- lavaan.survey(fitQ, survey.design = des, estimator="ML")
Fit <- rbind(fitMeasures(fitL)[c('chisq','df',"cfi",'tli', 'rmsea')], 
             fitMeasures(fitQ)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
anova(fitL, fitQ)
summary(fitQ, standardized=TRUE)
#By Group
mG_quadratic <- Model_Fun(name="FUTURE.PROS", params = list(i=T,s=T,q=T,ig=T,sg=T,qg=T))
mG_quadratic <- gsub("group", "group.x", mG_quadratic)
fitQG <- growth(mG_quadratic, data=dmatch, missing="fiml")
fitQG <- lavaan.survey(fitQG, survey.design = des, estimator="ML")
Fit <- rbind(Fit,  fitMeasures(fitQG)[c('chisq','df',"cfi",'tli', 'rmsea')])
Fit
summary(fitQG, standardized=TRUE)
##plot
paras<- parameterEstimates(fitQG)[c(44,13,20,21:23),"est"]
FunPlot_Multi2(model=fitQG, data=dmatch, main="Future Prospects",var="FUTURE.PROS", paras)
```

